Should I Airbnb My Property?

### 2017-03-08 abb_Writeup.Rmd
review - sensitivity analysis

To test if the above results are influence by the assumptions we have made throughout this analysis, we perform two sets of sensitivity tests.  First, we examine how the results changes if potential revenue, not observed revenue, of the short-term properties is used in the comparisons.  Second, we categorize the hosts, split the data and re-analyze the performance of properties based on hypothetical motivations and actions of the hosts.

@@

The above uses the observed^[Extrapolated out in cases of Airbnb properties on the market for less than a year] revenue in the comparisons with the imputed long-term revenue.  Doing so may underestimate the possible revenue -- what may matter most to the hypothetical buyer in our scenario -- due to the treatment of blocked days.  If we, instead, assume that blocked days could be leased out on the short-term market at the same overall occupancy rate as the unblocked (observed) days we can produce a potential revenue figure. Note that it is possible (likely?) that the large oversupply of available units onto the market that would happen if no properties blocked their units would decrease both prices and occupancy rates.  We do not attempt to correct for this likelihood, rather we offer these figures as high end estimates.

Using the potential revenue, 53.3% of all properties are short-term preferable.  Breaking that down by type and by geographic submarket shows similar finding to the actual revenue analysis above.  Apartments have a higher short-term preference than do Houses and City-Core properties dominate the other geographic regions.  One notable difference here is that the Rural properties perform rather poorly under a potential revenue scenario, relatively speaking, suggesting that they may be operating near their peak in the current condition.

@@

Examining the SVM heat maps for the entire metro region under the Actual revenue (left panel, Figure 19) and the Potential revenue (right panel) show the large increase in possible occupancy rate and nightly rate combination that lead to short-term preference.  Under a potential revenue scenario, a median occupancy rate and median nightly rate is results in greater than 50\% chance of short-term preference.  Even with very low nightly rates (10th percentile), an occupancy rate around the 65th percentile will achieve short-term prefernce likelihood.

@@

Our final test of the potential revenue is to re-run the logistic regression models using short-term preference in the potential case as the dependent variable. Generally speaking, the results from the new series of models (Table 6) are qualitatively similar to those derived in teh Actual revenue case (Table 4).  The overall magnitude of the coefficient for House (vs Apartments) is lowered in the Potential case, suggesting that houses may be blocked off more that Apartments.  The product types (bed/bath) show similar trends, though with larger discounts for the 2 Bed / 1 Bath and 2 Bed/2 Bath units.  Likewise the geographic submarkets and host policies show similar effects in terms of direction with some slight deviations in magnitude.

@@

Overall, the choice of revenue type, Actual or Potential, does affect some of the conclusions to be derived from the results above.  As expected, under the Potential case there are more properties (50\% vs 27\%) that exhibit short-term preference.  Under the scenario, the possibilities of increasing revenue by moving from long to short term leasing appear much better.  Again, however, we have not attempted to measure or estimate how this flood of properties onto the market (due to the elimation of blocked days) would impact overall market dynamics.  Suffice it to say that it would like drive down occupancies and maybe price, therefore leading to a decrease in short-term preference likelihood.

In explaining the drivers of short-term preference, the logistic regression models do not change materially with the switch in dependent variable.  This suggests that the underlying drivers of short-term preference remain relatively constant regardless of the choice of revenue estimate, only the base probability of short-term preference increases.

@@

As hinted at above in the Actual versus Potential comparison and as indicated in the logistic regression models in Tables 5 and 6, host behavior may be an important determinant of achieving short-term preference.  Additionally, as the difference between extrapolated and potential revenue suggests, many short-term hosts are not listing their properties full-time.  There may be, therefore, different types of hosts operating in the Airbnb market. from anecdotal evidence, there are at least two types of hosts in the market: 1) Those that lease out their properties all the time in the name of profit maximization (Profit Seekers); and 2) Those that lease out their property when very infrequently, for example when they are away on vacation (Opportunistic Sharers).

If these are the dominant two forms of hosts in the market, then a distribution of properties by the percent of time which they are blocked shoudl be very bi-modal with profit seekers on the left and opportunistic sharers on the right. The left hand panel of figure 9 shows that this breakdown hold somewhat, with peaks of hosts at each end, but that there are also a considerable number of hosts who block their properties some of the time.

There is also a third potential host type, that of the host who utilizies multiple platforms, of which Airbnb may just be one, to list their short-term property. A potential method to identify these hosts is to look at the number of individual periods of blocked days.  A host using multiple platforms will likely have many, small periods of blocked days on Airbnb as they will block out those days in which reservations are made on other platforms. We test to see if this is a meaningful metric by plotting the total count of blocked periods versus the percentage of blocked days in the right panel of Figure 9. We do find many host with a large number of blocked periods, but, unfortunately, no true pattern emerges.

@@

  host.block.freq <- ggplot(str.data, aes(x=block.rate)) + 
  geom_density(fill=str.col, color=str.col) +
  ylab('Frequency\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0,1,.2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Host Block Rate Frequency\n') +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())
```

@@

```{r plot_hostblock, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.caption='Host Blocking Tendencies', fig.lp='hostblock'}

@@

Attempts to divide the hosts up with unsupervised^[As there are no official host types we could not run more reliable supervised models.] clustering algorithms -- including the use of more variables -- failed to produce robust categories.  As a more subjective approach, we have divided the hosts up into four categories based on the original two variables considered: block percentage and number of blocked periods. Profit seekers are those host who have blocked percentages less than 25\% of the time.  Opportunistic Sharers have blocked their properties at least 75\% of the time.  Within those two extremes, we consider users who have at least 12 blocked periods during a year to be 'Multi-Platform Users' and the remaining set to be 'Unknown'.  See Figure 10 for a breakdown of these group by the two variables.

@@

```{r plot_hosttypes, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.caption='Host Categories', fig.lp='hosttypes'}

@@

Profit Seekers make up 43\% of the hosts, with Opportunistic Sharers capturing 18\% and Multi-Platform Users 10\%.  The remainder 'Unknown' category contains just over 28\% of all properties. The classification of these hosts is not convincing enough to base our entire analytical process on so we will use these breakdowns to add context to the results and to test the sensitivity of our results to the assumptions underlying the extrapolated and potential revenue calculations.

```{r plot_hostbar, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=4, fig.caption='Host Type Counts', fig.lp='hostbar'}

host.type.bar <- ggplot(str.data, aes(host.type, fill=host.type)) +
  geom_bar() +
  scale_fill_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9]))+
  theme(legend.position='none') +
  ylab('Frequency\n') + 
  xlab('\nHost Type') +
  ggtitle('Short Term Host Types') + 
  theme(plot.title = element_text(hjust = 0.5))

@@

Looking at the differences in short-term preference under both revenue cases, we see large variation in short-term preference between host types in the Actual case and a much smaller variation in the Potential case, as expected. There is little change for the 'Profit Seekers' as most of these hosts already operating with few blocked days.  For the other three sets, considerable gains could be had with a more aggresive leasing approach. As the potential figures are somewhat similar here, we will only examine the differences in the Actual number going forward.

@@

Table 9 compares logistic models results from three of the host types.  The Opportunistic Sharers show no instances of short-term preference and therefore could not be estimated.  The left hand column, that of the Profit-Seekers, is the most representative of a sample of property owners that are seeking to maximize revenue.  Overall, the coefficients from this sub-sample match up well to the results from the entire sample shown in Table 5.  In short, the conclusions are the same: small, apartments in the central city with short minimal stays, inflexible cancelation policies and a high guest limit offer the highest probability of short-term preference.  The models for the multi-platform users (middle panel) and the Unknowns (right) are not as clear as the Profit-Seekers but do suggest overall similar dynamics.

@@

 xh.full <- xtable(hfull.table,
                  caption='Logistic Modeling Results\n')
 align(xh.full) <- 'lllll'

>>>>

To test if the above results are influence by the assumptions we have made throughout this analysis, we perform two sets of sensitivity tests.  First, we examine how the results change if Potential revenue, not Actual revenue, of the short-term properties is used in the comparisons.  Second, we categorize the hosts, split the data and re-analyze the performance of properties based on hypothetical motivations and actions of the hosts.

@@

The above uses the Actual^[Extrapolated out in cases of Airbnb properties on the market for less than a year] revenue in the comparisons with the imputed long-term revenue.  Doing so may underestimate the possible revenue -- what may matter most to the hypothetical buyer in our scenario -- due to the treatment of blocked days.  If we, instead, assume that blocked days could be leased out on the short-term market at the same overall occupancy rate as the unblocked (observed) days we can produce a potential revenue figure. Note that it is possible (likely?) that the large oversupply of available units onto the market that would happen if no properties blocked their units would decrease both prices and occupancy rates.  We do not attempt to correct for this likelihood, rather we offer these figures as high end estimates.

Using the potential revenue, 53.3% of all properties are short-term preferable.  Breaking that down by type and by geographic submarket (Table 5) shows similar findings, in terms of relative comparativeness, to the actual revenue analysis above.  Apartments have a higher short-term preference than do Houses and City-Core properties dominate the other geographic regions.  One notable difference here is that the Rural properties perform rather poorly under a potential revenue scenario, relatively speaking, suggesting that they may be operating near their peak in the current condition.

@@

Examining the SVM heat maps for the entire metro region under the Actual revenue (left panel, \autoref{potlog}) and the Potential revenue (right panel) show the large increase in possible occupancy rate and nightly rate combinations that lead to short-term preference.  Under a potential revenue scenario, a median occupancy rate and median nightly rate is results in greater than 50\% chance of short-term preference.  Even with very low nightly rates (10th percentile), an occupancy rate around the 65th percentile will achieve short-term preference likelihood in the potential case.

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='\\label{potlog}Logistic Model Results of Short Term Preference', fig.lp='xxx'}

@@

Our final test of the potential revenue case is to re-run the logistic regression models using short-term preference in the potential case as the dependent variable. Generally speaking, the results from the new series of models (Table 6) are qualitatively similar to those derived in the Actual revenue case (Table 4).  The overall magnitude of the coefficient for House (vs Apartments) is lowered in the Potential case, suggesting that houses may be blocked off more that Apartments.  The product types (bed/bath) show similar trends, though with larger discounts for the 2 Bed / 1 Bath and 2 Bed/2 Bath units.  Likewise the geographic submarkets and host policies show similar effects in terms of direction of impact with some slight deviations in magnitude.

@@

Overall, the choice of revenue type, Actual or Potential, does affect some of the conclusions to be derived from the results above.  As expected, under the Potential case there are more properties (50\% vs 27\%) that exhibit short-term preference.  Under the scenario, the possibilities of increasing revenue by moving from long to short term leasing appear much better.  Again, however, we have not attempted to measure or estimate how this hypothetical flood of properties onto the market (due to the elimation of blocked days) would impact overall market dynamics.  Suffice it to say that it would like drive down occupancies and maybe price, therefore leading to a decrease in short-term preference likelihood.

In explaining the drivers of short-term preference, the logistic regression models do not change materially with the switch in dependent variable.  This suggests that the underlying drivers of short-term preference remain relatively constant regardless of the choice of revenue estimate. Only the base probability of short-term preference increases in the the potential revenue case

@@

As hinted at above in the Actual versus Potential comparison and as indicated in the logistic regression models in Tables 5 and 6, host behavior may be an important determinant of achieving short-term preference.  Additionally, as the difference between extrapolated and potential revenue suggests, many short-term hosts are not listing their properties full-time.  There may be, therefore, different types of hosts operating in the Airbnb market. From anecdotal evidence there are at least two types of hosts in the market: 1) Those that lease out their properties all the time in the name of profit maximization (Profit Seekers); and 2) Those that lease out their property very infrequently, for example when they are away on vacation (Opportunistic Sharers).

If these are the dominant two forms of hosts in the market, then a distribution of properties by the percent of time which they are blocked should be highly bi-modal with profit seekers on the left and opportunistic sharers on the right. The left hand panel of \autoref{hostblock} shows that this breakdown holds somewhat, with peaks of hosts at each end, but that there are also a considerable number of hosts who block their properties some (25\% to 75\%) of the time.

There is also a third potential host type, that of the host who utilizies multiple platforms to list their short-term property, of which Airbnb may just be one. A potential method to identify these hosts is to look at the number of individual periods of blocked days.  A host using multiple platforms will likely have many small periods of blocked days on Airbnb as they will block out those days in which reservations are made on other platforms. We test to see if this is a meaningful metric by plotting the total count of blocked periods versus the percentage of blocked days in the right panel of \autoref{hostblock}. We do find many hosts with a large number of blocked periods, but, unfortunately, no true pattern emerges.

@@

  host.block.freq <- ggplot(str.data, aes(x=block.rate)) + 
  geom_density(fill=str.col, color=str.col) +
  ylab('Frequency\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0,1,.2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    #ggtitle('Host Block Rate Frequency\n') +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank())
```

@@

```{r plot_hostblock, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.caption='\\label{hostblock}Host Blocking Tendencies', fig.lp='hostblock'}

@@

Attempts to divide the hosts up with unsupervised^[As there are no official host types we could not run more reliable supervised models.] clustering algorithms -- including the use of more variables -- failed to produce robust categories of hosts. As a more subjective, but implementable, approach we have divided the hosts up into four categories based on the original two variables considered: block percentage and number of blocked periods. Profit seekers are those host who have blocked percentages less than 25\% of the time.  Opportunistic Sharers have blocked their properties at least 75\% of the time.  Within those two extremes, we consider users who have at least 12 blocked periods during a year to be 'Multi-Platform Users' and the remaining set to be 'Unknown'. Profit Seekers make up 43\% of the hosts, with Opportunistic Sharers capturing 18\% and Multi-Platform Users 10\%.  The remainder 'Unknown' category contains just over 28\% of all properties. See \autoref{hostcat} and \autoref{htc} for a breakdown of these group by the two variables.

@@

```{r plot_hosttypes, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.caption='\\label{hostcat}Host Categories', fig.lp='hosttypes'}

@@

```{r plot_hostbar, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3, fig.width=4 fig.caption='\\label{htc}Host Type Counts', fig.lp='hostbar'}

host.type.bar <- ggplot(str.data, aes(host.type, fill=host.type)) +
  geom_bar() +
  scale_fill_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9]))+
  theme(legend.position='none') +
  ylab('Frequency\n') + 
  xlab('\nHost Type') +
  #ggtitle('Short Term Host Types') + 
  theme(plot.title = element_text(hjust = 0.5))

@@

Looking at the differences in short-term preference under both revenue cases, we see large variation in short-term preference between host types in the Actual case and a much smaller variation in the Potential case, as expected. There is little change for the 'Profit Seekers' as most of these hosts are already operating with few blocked days.  For the other three sets, considerable gains could be had with a more aggresive leasing approach. As the potential figures are somewhat similar here, we will only examine the differences for host types in the Actual number going forward.

@@

Table 8 compares logistic models results from three of the host types.  The Opportunistic Sharers show no instances of short-term preference and therefore could not be estimated.  The left hand column, that of the Profit-Seekers, is the most representative of a sample of property owners that are seeking to maximize revenue.  Overall, the coefficients from this sub-sample match up well to the results from the entire sample shown in Table 5.  In short, the conclusions are the same: small apartments in the central city with short minimal stays, inflexible cancelation policies and a high guest limit offer the highest probability of short-term preference.  The models for the multi-platform users (middle panel) and the Unknowns (right) are not as clear as the Profit-Seekers but do suggest overall similar dynamics.

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='\\label{logpot}Logistic Model Results of Potential Short Term Preference'}

@@

 xh.full <- xtable(hfull.table,
                  caption='Logistic Model Results of Potential Short-Term Preference\n')
 align(xh.full) <- 'lllll'
